---
title: "This is how the decline of the SPD sounds"
output: github_document
number_sections: yes
---

`December 13, 2024`

Germany's Social Democratic Party (SPD) still leads the current government under Chancellor Olaf Scholz – but his governing coalition has fallen apart and he's called for snap elections. The SPD's poll numbers have been declining for years. We turned the figures into sound – the lower the note, the worse the result.

*This story is inspired by the 2018 project ["Der Sound zum tiefen Fall der SPD"](https://interaktiv.morgenpost.de/spd-absturz-sound/) by Funke Interaktiv for Berliner Morgenpost.*

*Thank you to Moritz Klack, André Pätzold, Marie-Louise Timcke, Julius Tröger and David Wendler for their great work and for making the [code behind their project](https://interaktiv.morgenpost.de/spd-absturz-sound/data/spd-absturz-sound-methodik.html) publicly available.*

See the final result on our Instagram channel: [\@dw_news](https://www.instagram.com/dwnews)

**Story by:** [Kira Schacht](https://twitter.com/daten_drang) and [Dustin Hemmerlein](https://de.linkedin.com/in/dustin-hemmerlein-4671a8173)

## Analysis 

You can find the code behind this analysis in the R Markdown file `SPD-polls-analysis.Rmd`. You will need the programming language R to run it.

```{r setup, include=FALSE}
setwd("/Users/SchachtK/Library/CloudStorage/OneDrive-DeutscheWelle/2412 SPD decline/github/spd-decline")

library("needs")
needs(tidyverse, rvest, janitor)

knitr::opts_chunk$set(echo = FALSE)
```

### Scrape data

To analyze polling results, we extract data from [wahlrecht.de](https://www.wahlrecht.de/umfragen/index.htm), which has been documenting polling results in Germany going back to the 1990s.

```{r Make list of links to scrape}
#links to each pollster
tmp = read_html("http://www.wahlrecht.de/umfragen/") %>%
      html_nodes("th.in > a") %>% 
      html_attr("href") %>% #get links from html structure
      paste0("http://www.wahlrecht.de/umfragen/", .) #format
#links to past legislation periods for each of these pollsters
tmp2 = lapply(tmp, function(l) {
  read_html(l) %>% html_nodes("p.navi > a") %>% html_attr("href")
  }) %>% #get links from html structure
  unlist %>% #concatenate
  paste0("http://www.wahlrecht.de/umfragen/", .) %>% #format
  `[`(grepl("[0-9]",.)) #filter out links that don't lead to previous time periods

#bind both link lists together
l = c(tmp, tmp2)

rm(tmp,tmp2) #remove temporary variables

```

```{r Scrape and clean data, warning=FALSE}

#Create empty list to fill with data
wahlrecht = vector("list", length = length(l))

#Loop through each sub-page
for(i in 1:length(l)){
  #read source code
  tmp = read_html(l[i])
  #get pollster name
  pollster = tmp %>% html_node("h1") %>% html_text()
  
  wahlrecht[[i]] = tmp %>%
    #extract table node from html
    html_node("table.wilko") %>% 
    #convert to data frame
    html_table(convert = FALSE, trim = TRUE) %>% 
    #use first row as headers
    row_to_names(row_number = 1, remove_row = TRUE) %>% 
    clean_names() %>% 
    #append pollster name
    cbind(pollster = pollster)
      
};rm(i, tmp, pollster)

#clean resulting data frame
wahlrecht = wahlrecht %>%
  bind_rows() %>% #convert to one data frame
  select(1,3:8,15) %>% #only keep date, current parliament parties, and pollster name
  rename(datum = x) %>% 
  mutate(datum = gsub("\\?\\?","01",datum)) %>% #replace one instance of unknown date with first of the month
  filter(grepl("^[0-9]", datum)) %>% #filter out rows without poll date (empty rows or meta info)
  mutate(across(2:7, ~ gsub("[ \\%–]","",.x) )) %>% #format results as numbers
  mutate(across(2:7, ~ gsub(",",".",.x) )) %>% 
  mutate(across(2:7, as.numeric )) %>% 
  mutate(datum = convert_to_date(datum, character_fun = lubridate::dmy)) %>% 
  filter(!is.na(spd)) #remove two rows without polling data

save(wahlrecht, file = "data/poll_data.RData")
```

This is an excerpt of the resulting dataset:

```{r}
wahlrecht %>% arrange(desc(datum)) %>% head()
```

And this is what all of these polling results look like over time:

```{r warning=FALSE}
ggplot(wahlrecht %>% filter(datum >= as.Date("2005-10-01")),
       aes(datum, spd)) +
  geom_point(color="#DB4240", alpha=0.09, stroke=0, linewidth=2) + 
  labs(title = "SPD polling results over time, in %", x = "", y = "") +
  theme_minimal()
```

### Calculate smoothed average line

We include data from representative surveys from 8 different pollsters in this analysis, which means there might be multiple results with slight statistical variations for the same time period.

In order to show an average of these polls over time, we use a local regression ([LOESS-smoothing](https://web.archive.org/web/20020416060643/https://www.itl.nist.gov/div898/handbook//pmd/section1/pmd144.htm)) algorithm.

For any given point in time, it considers the closest 2,5 percent of survey values in our dataset and calculates a weighted average of those. The closer to the group average a value is, the more it factors into the calculation. This limits the effect of extreme outliers on our estimated average.

```{r Calculate local regression}
#fit loess smoothing (local regression) model with span of 2.5%
model = loess(wahlrecht$spd ~ as.numeric(wahlrecht$datum), span=0.025)
```

In order to turn the polling average into distinct sounds, we show one data point per month instead of a continuous daily curve. We show the last 20 years, starting with the first month after the federal election in 2005.

```{r Calculate one prediction per month}
#round to monthly predictions for the sonification 
months = seq(as.Date('2005-10-01'), as.Date('2024-12-01'), by="months")
predictions = data.frame(date = months,
                    value = predict(model, as.numeric(months)) %>% round)
```

This is what the finished chart looks like:

```{r}
ggplot(wahlrecht %>% filter(datum >= as.Date("2005-10-01")),
       aes(datum, spd)) +
  geom_point(color="#DB4240", alpha=0.09, stroke=0, size=2) + 
  geom_line(data = predictions, aes(date, value), color = "#DB4240", size=1)+
  labs(title = "SPD polling results over time, in %", x = "", y = "") +
  theme_minimal()
```

### Sonify data

To turn the monthly local averages into sound, we use the free tool [DataSonifyer](https://studio.datasonifyer.de/).

If you want to see which presets we used, you can import our configuration file `data/spd-polls-DataSonifyerExport.json` into DataSonifyer with the "Import preset" button at the bottom of the tool's page.

```{r Export data}
predictions %>%
  write.csv("data/spd-polls-monthly-loess.csv",
            row.names = F, na = "", quote = F)
```
